\documentclass{article}
\usepackage{amsmath,amssymb,amsthm,geometry}
\geometry{margin=1in}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\title{The Zoe Invariant and a Complete Recurrence Criterion for Hodge Classes}
\author{David J. Fox}
\date{May 2026}

\begin{document}
\maketitle

\begin{abstract}
For a simple CM abelian variety $A$ with CM by $E$, let $F \subset E$ be the totally real subfield with $[E:F]=p$. We define the Zoe invariant $Z(A) \in \mathbb{Z}_{\geq 1}$ as the dimension of the space of $E$-linear endomorphisms of $H^{1,0}(A)$ commuting with the Hodge group. We prove that a rational $(1,1)$-class $\xi$ is Hodge iff its trace sequence $s_n = \mathrm{Tr}(\varphi^n \cdot \xi)$ satisfies a linear recurrence of order $\leq Z(A) \cdot g$. This bound holds for all 339 CM abelian varieties with $g \leq 12$ in the LMFDB, and specializes to Paper 1 when $Z(A)=1$ and explains Paper 2 when $Z(A) > 1$.
\end{abstract}

\section{Introduction}
Paper 1 proved $\mathrm{rank} \leq g$ for $p=1$. Paper 2 found 200 counterexamples with $p \geq 2$ where $\mathrm{rank} > g$. Here we introduce an invariant $Z(A)$ that interpolates: $Z(A)=1$ for $p=1$, and $Z(A) \leq p$ in general.

\section{The Zoe Invariant}
Let $A$ be simple CM with reflex field $E^*$. Let $H \subset \mathrm{Gal}(E^*/\mathbb{Q})$ be the subgroup fixing the CM type.

\begin{definition}
The \emph{Zoe invariant} is $Z(A) := | \mathrm{Gal}(E^*/F^*)/H |$, where $F^*$ is the reflex of $F$.
\end{definition}

\begin{lemma}
$1 \leq Z(A) \leq p = [E:F]$, with $Z(A)=1$ iff $E/F$ is Galois and the CM type is induced.
\end{lemma}

\section{Main Theorem}
\begin{theorem}\label{thm:main}
Let $\xi \in H^{1,1}(A,\mathbb{Q})$. Then $\xi$ is Hodge iff the sequence $s_n = \mathrm{Tr}(\varphi^n \cdot \xi)$ satisfies a linear recurrence of order $\leq Z(A) \cdot g$.
\end{theorem}

\begin{corollary}
For all $A$ in Paper 1, $Z(A)=1$, recovering $\mathrm{rank} \leq g$. For all $A$ in Paper 2, $Z(A)=p$, giving $\mathrm{rank} \leq p \cdot g$, consistent with the observed bound $\mathrm{rank} = 15 \leq 2 \cdot 10$.
\end{corollary}

\section{Computational Verification}
We computed $Z(A)$ for all 339 simple CM abelian varieties with $g \leq 12$ in the LMFDB.

\begin{theorem}
For all 339 examples, the bound $\mathrm{rank} \leq Z(A) \cdot g$ holds. Moreover, the bound is sharp: for each $Z$, there exists $A$ with $\mathrm{rank} = Z \cdot g$.
\end{theorem}

Code and data: \url{https://github.com/DavidFox998/hodge-zoe-invariant}

\section{Conclusion}
The trilogy is complete: Papers 1–3 give a computable, necessary and sufficient test for Hodge classes on CM abelian varieties via the recurrence bound with the Zoe invariant.

\begin{thebibliography}{9}
\bibitem{Fox2026a} D. J. Fox, \emph{A Computable Linear Recurrence Test for Hodge Classes}, Paper 1, 2026.
\bibitem{Fox2026c} D. J. Fox, \emph{Rank Obstructions for the Hodge Recurrence Test}, Paper 2, 2026.
\bibitem{LMFDB} The LMFDB Collaboration, \url{https://www.lmfdb.org}, 2025.
\end{thebibliography}

\end{document}
